Abundance/diversity GAM

### 2020-09-15 analyses/abund_div_GAM.Rmd
refined GDM analysis

### Call GAMs
##### My argument for not including site as a random effect is that it would create identifiability issues with the elevation variable that I am actally interested in. This will probably cause the p-values to be underestimated, maybe to the point of being uninformative. But I don't think I care. My goal is not to establish inferentially that time and elevation affect BB abundance; I want to say how they do. And I don't want to obscure the "how" for the sake of inferential scruples.
#### Floral resources
##### After trying a lot of things, I'm pretty sure the only option I have for fitting a good GAM is using log-transformation. The question that remains is how to treat the concept of preference. My "raw" preference-weighted abundance gives our measure of preference a lot of influence. If we trust it, this may be most informative, and it allows the greatest variation in floral resources between BB species. But if there is anything misleading in our preference estimates, using the "raw" measures will do nothing to mitigate this. An approach that gives us a "softer" notion of preference is to take the sqrt of the raw prference before multiplying floral abundance by it. An argument for this approach is that it retains most of the meaning of the raw preference data while dampening its effect, which may help avoid spurious hotspots based on overestimated preference. Finally, an approach that says "we don't really trust our data very much" is to treat preference as binary: visited species get a 1, non-visited species get a 0, and all the "weighting" of floral resource abundance means is dropping floral species that a given BB species did not visit. In all these cases, though, the distribution of weighted floral abundance is strongly logarithmic (or exponential or something), so a log transformation is necessary. I tried softer methods, e.g. sqrt-transformation, Tweedie family, etc., but the problem with all these is that the variance fans out like crazy as you move up to the higher end of the distribution. Happily, it makes little difference which of these options we choose, as the resulting plots lead to the same interpretation.
```{r message=FALSE}
# These are the distributions of my response variable options 
ggplot(fl_abund_wt, aes(fl.abund.wt.bin)) +
  geom_histogram()

@@

check.gamViz(weighted_fl_abundance.2010)
#weighted_fl_abundance.2010.sum <- summary(weighted_fl_abundance.2010)
#print(plot(weighted_fl_abundance.2010, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2010.ck <- check2D(weighted_fl_abundance.2010, x1 = "yday", x2 = "elev.mean")

@@

check.gamViz(weighted_fl_abundance.2010.sqrt)
#weighted_fl_abundance.2010.sum.sqrt <- summary(weighted_fl_abundance.2010.sqrt)
#print(plot(weighted_fl_abundance.2010.sqrt, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2010.ck.sqrt <- check2D(weighted_fl_abundance.2010.sqrt, x1 = "yday", x2 = "elev.mean")

@@

check.gamViz(weighted_fl_abundance.2010.bin)
#weighted_fl_abundance.2010.sum.bin <- summary(weighted_fl_abundance.2010.bin)
#print(plot(weighted_fl_abundance.2010.bin, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2010.ck.bin <- check2D(weighted_fl_abundance.2010.bin, x1 = "yday", x2 = "elev.mean")

@@

check.gamViz(weighted_fl_abundance.2011)
#weighted_fl_abundance.2011.sum <- summary(weighted_fl_abundance.2011)
#print(plot(weighted_fl_abundance.2011, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2011.ck <- check2D(weighted_fl_abundance.2011, x1 = "yday", x2 = "elev.mean")

@@

check.gamViz(weighted_fl_abundance.2011.sqrt)
#weighted_fl_abundance.2011.sum.sqrt <- summary(weighted_fl_abundance.2011.sqrt)
#print(plot(weighted_fl_abundance.2011.sqrt, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2011.ck.sqrt <- check2D(weighted_fl_abundance.2011.sqrt, x1 = "yday", x2 = "elev.mean")

@@

check.gamViz(weighted_fl_abundance.2011.bin)
#weighted_fl_abundance.2011.sum.bin <- summary(weighted_fl_abundance.2011.bin)
#print(plot(weighted_fl_abundance.2011.bin, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2011.ck.bin <- check2D(weighted_fl_abundance.2011.bin, x1 = "yday", x2 = "elev.mean")

@@

check.gamViz(weighted_fl_abundance.2012)
#weighted_fl_abundance.2012.sum <- summary(weighted_fl_abundance.2012)
#print(plot(weighted_fl_abundance.2012, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2012.ck <- check2D(weighted_fl_abundance.2012, x1 = "yday", x2 = "elev.mean")

@@

check.gamViz(weighted_fl_abundance.2012.sqrt)
#weighted_fl_abundance.2012.sum.sqrt <- summary(weighted_fl_abundance.2012.sqrt)
#print(plot(weighted_fl_abundance.2012.sqrt, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2012.ck.sqrt <- check2D(weighted_fl_abundance.2012.sqrt, x1 = "yday", x2 = "elev.mean")

@@

check.gamViz(weighted_fl_abundance.2012.bin)
#weighted_fl_abundance.2012.sum.bin <- summary(weighted_fl_abundance.2012.bin)
#print(plot(weighted_fl_abundance.2012.bin, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2012.ck.bin <- check2D(weighted_fl_abundance.2012.bin, x1 = "yday", x2 = "elev.mean")
```

@@

check.gamViz(bb_abund.2010)
#bb_abund.2010.sum <- summary(bb_abund.2010)
#print(plot(bb_abund.2010, allTerms = TRUE), pages = 1)
#bb_abund.2010.ck <- check2D(bb_abund.2010, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals

@@

check.gamViz(bb_abund.2011)
#bb_abund.2011.sum <- summary(bb_abund.2011)
#print(plot(bb_abund.2011, allTerms = TRUE), pages = 1)
#bb_abund.2011.ck <- check2D(bb_abund.2011, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals

@@

check.gamViz(bb_abund.2012)
#bb_abund.2012.sum <- summary(bb_abund.2012)
#print(plot(bb_abund.2012, allTerms = TRUE), pages = 1)
#bb_abund.2012.ck <- check2D(bb_abund.2012, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals
```

@@

predict_pool.FL <- bind_rows(predict2010.FL, predict2011.FL, predict2012.FL) %>%
  mutate(date = as_date(yday))
predict_pool.FL.sqrt <- bind_rows(predict2010.FL.sqrt, predict2011.FL.sqrt, predict2012.FL.sqrt) %>%
  mutate(date = as_date(yday))
predict_pool.FL.bin <- bind_rows(predict2010.FL.bin, predict2011.FL.bin, predict2012.FL.bin) %>%
  mutate(date = as_date(yday))

predict2010.BB <- predict_func(bb_abund.2010, newdata2010.BB, 2010)
predict2011.BB <- predict_func(bb_abund.2011, newdata2011.BB, 2011)
predict2012.BB <- predict_func(bb_abund.2012, newdata2012.BB, 2012)
predict_pool.BB <- bind_rows(predict2010.BB, predict2011.BB, predict2012.BB) %>%
  mutate(date = as_date(yday))

# Subset predictions
predict_major.FL <- filter(predict_pool.FL, 
                           bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf")) %>%
  mutate(bb.sp = case_when(
    bb.sp == "bss" ~ "B. sensu-strictu",
    bb.sp == "pasc" ~ "B. pascuorum",
    bb.sp == "prat" ~ "B. pratorum",
    bb.sp == "soro" ~ "B. soroensis",
    bb.sp == "wurf" ~ "B. wurflenii"
  ))
predict_major.FL.sqrt <- filter(predict_pool.FL.sqrt, 
                                bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf")) %>%
  mutate(bb.sp = case_when(
    bb.sp == "bss" ~ "B. sensu-strictu",
    bb.sp == "pasc" ~ "B. pascuorum",
    bb.sp == "prat" ~ "B. pratorum",
    bb.sp == "soro" ~ "B. soroensis",
    bb.sp == "wurf" ~ "B. wurflenii"
  ))
predict_major.FL.bin <- filter(predict_pool.FL.bin, 
                               bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf")) %>%
  mutate(bb.sp = case_when(
    bb.sp == "bss" ~ "B. sensu-strictu",
    bb.sp == "pasc" ~ "B. pascuorum",
    bb.sp == "prat" ~ "B. pratorum",
    bb.sp == "soro" ~ "B. soroensis",
    bb.sp == "wurf" ~ "B. wurflenii"
  ))

predict_highelev.FL <- filter(predict_pool.FL, 
                              bb.sp %in% c("pyre", "mend", "mont")) %>%
  mutate(bb.sp = case_when(
    bb.sp == "pyre" ~ "B. pyrenaeus",
    bb.sp == "mend" ~ "B. mendax",
    bb.sp == "mont" ~ "B. monticola"
  ))
predict_highelev.FL.sqrt <- filter(predict_pool.FL.sqrt, 
                                   bb.sp %in% c("pyre", "mend", "mont")) %>%
  mutate(bb.sp = case_when(
    bb.sp == "pyre" ~ "B. pyrenaeus",
    bb.sp == "mend" ~ "B. mendax",
    bb.sp == "mont" ~ "B. monticola"
  ))
predict_highelev.FL.bin <- filter(predict_pool.FL.bin, 
                                  bb.sp %in% c("pyre", "mend", "mont")) %>%
  mutate(bb.sp = case_when(
    bb.sp == "pyre" ~ "B. pyrenaeus",
    bb.sp == "mend" ~ "B. mendax",
    bb.sp == "mont" ~ "B. monticola"
  ))

predict_minor.FL <- filter(predict_pool.FL, 
                           bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci")) %>%
  mutate(bb.sp = case_when(
    bb.sp == "gers" ~ "B. gerstaeckeri",
    bb.sp == "hort" ~ "B. hortorum",
    bb.sp == "hypn" ~ "B. hypnorum",
    bb.sp == "jone" ~ "B. jonellus",
    bb.sp == "lapi" ~ "B. lapidarius",
    bb.sp == "muci" ~ "B. mucidus"
  ))
predict_minor.FL.sqrt <- filter(predict_pool.FL.sqrt, 
                                bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci")) %>% 
  mutate(bb.sp = case_when(
    bb.sp == "gers" ~ "B. gerstaeckeri",
    bb.sp == "hort" ~ "B. hortorum",
    bb.sp == "hypn" ~ "B. hypnorum",
    bb.sp == "jone" ~ "B. jonellus",
    bb.sp == "lapi" ~ "B. lapidarius",
    bb.sp == "muci" ~ "B. mucidus"
  ))
predict_minor.FL.bin <- filter(predict_pool.FL.bin, 
                               bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci")) %>% 
  mutate(bb.sp = case_when(
    bb.sp == "gers" ~ "B. gerstaeckeri",
    bb.sp == "hort" ~ "B. hortorum",
    bb.sp == "hypn" ~ "B. hypnorum",
    bb.sp == "jone" ~ "B. jonellus",
    bb.sp == "lapi" ~ "B. lapidarius",
    bb.sp == "muci" ~ "B. mucidus"
  ))

predict_major.BB <- filter(predict_pool.BB, 
                           bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf")) %>%
  mutate(bb.sp = case_when(
    bb.sp == "bss" ~ "B. sensu-strictu",
    bb.sp == "pasc" ~ "B. pascuorum",
    bb.sp == "prat" ~ "B. pratorum",
    bb.sp == "soro" ~ "B. soroensis",
    bb.sp == "wurf" ~ "B. wurflenii"
  ))
predict_highelev.BB <- filter(predict_pool.BB, 
                              bb.sp %in% c("pyre", "mend", "mont")) %>%
  mutate(bb.sp = case_when(
    bb.sp == "pyre" ~ "B. pyrenaeus",
    bb.sp == "mend" ~ "B. mendax",
    bb.sp == "mont" ~ "B. monticola"
  ))
predict_minor.BB <- filter(predict_pool.BB, 
                           bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci")) %>%
  mutate(bb.sp = case_when(
    bb.sp == "gers" ~ "B. gerstaeckeri",
    bb.sp == "hort" ~ "B. hortorum",
    bb.sp == "hypn" ~ "B. hypnorum",
    bb.sp == "jone" ~ "B. jonellus",
    bb.sp == "lapi" ~ "B. lapidarius",
    bb.sp == "muci" ~ "B. mucidus"
  ))
```

### Visualize
```{r}
### FL GAMS
#### Major
ggplot(predict_major.FL, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(wt. abund.)") +
  facet_grid(year~bb.sp, ) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(strip.text.x = element_text(face = "italic"),
        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
ggsave("./output/GAM_major_FL.png")

ggplot(predict_major.FL.sqrt, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(wt. abund.)") +
  facet_grid(year~bb.sp) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(strip.text.x = element_text(face = "italic"),
        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
ggsave("./output/GAM_major_FL_sqrt.png")

ggplot(predict_major.FL.bin, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(wt. abund.)") +
  facet_grid(year~bb.sp) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(strip.text.x = element_text(face = "italic"),
        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
ggsave("./output/GAM_major_FL_bin.png")

#### High Elev
ggplot(predict_highelev.FL, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(wt. abund.)") +
  facet_grid(year~bb.sp) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(strip.text.x = element_text(face = "italic"),
        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
ggsave("./output/GAM_highelev_FL.png")

ggplot(predict_highelev.FL.sqrt, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  scale_fill_viridis_c(name = "log(wt. abund.)") +
  facet_grid(year~bb.sp) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(strip.text.x = element_text(face = "italic"),
        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
ggsave("./output/GAM_highelev_FL_sqrt.png")

ggplot(predict_highelev.FL.bin, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(wt. abund.)") +
  facet_grid(year~bb.sp) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(strip.text.x = element_text(face = "italic"),
        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
ggsave("./output/GAM_highelev_FL_bin.png")

#### Minor
ggplot(predict_minor.FL, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(wt. abund.)") +
  facet_grid(year~bb.sp) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(strip.text.x = element_text(face = "italic"),
        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
ggsave("./output/GAM_minor_FL.png")

ggplot(predict_minor.FL.sqrt, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(wt. abund.)") +
  facet_grid(year~bb.sp) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(strip.text.x = element_text(face = "italic"),
        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
ggsave("./output/GAM_minor_FL_sqrt.png")

ggplot(predict_minor.FL.bin, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(wt. abund.)") +
  facet_grid(year~bb.sp) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(strip.text.x = element_text(face = "italic"),
        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
ggsave("./output/GAM_minor_FL_bin.png")

### BB GAMs
ggplot(predict_major.BB, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = log(pred)), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "wt. abund.") +
  facet_grid(year~bb.sp) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(strip.text.x = element_text(face = "italic"),
        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
ggsave("./output/GAM_major_BB.png")

ggplot(predict_highelev.BB, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = log(pred)), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "wt. abund.") +
  facet_grid(year~bb.sp) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(strip.text.x = element_text(face = "italic"),
        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
ggsave("./output/GAM_highelev_BB.png")

>>>>

### Call GAMs
##### My argument for not including site as a random effect is that it would create identifiability issues with the elevation variable that I am actally interested in. This will probably cause the p-values to be underestimated, maybe to the point of being uninformative. But I don't think I care. My goal is not to establish inferentially that time and elevation affect BB abundance; I want to say how they do. And I don't want to obscure the "how" for the sake of inferential scruples.
#### Floral resources
##### After trying a lot of things, I'm pretty sure the only option I have for fitting a good GAM is using log-transformation. The question that remains is how to treat the concept of preference. My "raw" preference-weighted abundance gives our measure of preference a lot of influence. If we trust it, this may be most informative, and it allows the greatest variation in floral resources between BB species. But if there is anything misleading in our preference estimates, using the "raw" measures will do nothing to mitigate this. An approach that gives us a "softer" notion of preference is to take the sqrt of the raw prference before multiplying floral abundance by it. An argument for this approach is that it retains most of the meaning of the raw preference data while dampening its effect, which may help avoid spurious hotspots based on overestimated preference. Finally, an approach that says "we don't really trust our data very much" is to treat preference as binary: visited species get a 1, non-visited species get a 0, and all the "weighting" of floral resource abundance means is dropping floral species that a given BB species did not visit. In all these cases, though, the distribution of weighted floral abundance is strongly logarithmic (or exponential or something), so a log transformation is necessary. I tried softer methods, e.g. sqrt-transformation, Tweedie family, etc., but the problem with all these is that the variance fans out like crazy as you move up to the higher end of the distribution. Happily, it makes little difference which of these options we choose, as the resulting plots lead to the same interpretation.
```{r message=FALSE, warning=FALSE}
# These are the distributions of my response variable options 
ggplot(fl_abund_wt, aes(fl.abund.wt.bin)) +
  geom_histogram()

@@

#check.gamViz(weighted_fl_abundance.2010)
#weighted_fl_abundance.2010.sum <- summary(weighted_fl_abundance.2010)
#print(plot(weighted_fl_abundance.2010, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2010.ck <- check2D(weighted_fl_abundance.2010, x1 = "yday", x2 = "elev.mean")

@@

#check.gamViz(weighted_fl_abundance.2010.sqrt)
#weighted_fl_abundance.2010.sum.sqrt <- summary(weighted_fl_abundance.2010.sqrt)
#print(plot(weighted_fl_abundance.2010.sqrt, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2010.ck.sqrt <- check2D(weighted_fl_abundance.2010.sqrt, x1 = "yday", x2 = "elev.mean")

@@

#check.gamViz(weighted_fl_abundance.2010.bin)
#weighted_fl_abundance.2010.sum.bin <- summary(weighted_fl_abundance.2010.bin)
#print(plot(weighted_fl_abundance.2010.bin, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2010.ck.bin <- check2D(weighted_fl_abundance.2010.bin, x1 = "yday", x2 = "elev.mean")

@@

#check.gamViz(weighted_fl_abundance.2011)
#weighted_fl_abundance.2011.sum <- summary(weighted_fl_abundance.2011)
#print(plot(weighted_fl_abundance.2011, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2011.ck <- check2D(weighted_fl_abundance.2011, x1 = "yday", x2 = "elev.mean")

@@

#check.gamViz(weighted_fl_abundance.2011.sqrt)
#weighted_fl_abundance.2011.sum.sqrt <- summary(weighted_fl_abundance.2011.sqrt)
#print(plot(weighted_fl_abundance.2011.sqrt, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2011.ck.sqrt <- check2D(weighted_fl_abundance.2011.sqrt, x1 = "yday", x2 = "elev.mean")

@@

#check.gamViz(weighted_fl_abundance.2011.bin)
#weighted_fl_abundance.2011.sum.bin <- summary(weighted_fl_abundance.2011.bin)
#print(plot(weighted_fl_abundance.2011.bin, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2011.ck.bin <- check2D(weighted_fl_abundance.2011.bin, x1 = "yday", x2 = "elev.mean")

@@

#check.gamViz(weighted_fl_abundance.2012)
#weighted_fl_abundance.2012.sum <- summary(weighted_fl_abundance.2012)
#print(plot(weighted_fl_abundance.2012, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2012.ck <- check2D(weighted_fl_abundance.2012, x1 = "yday", x2 = "elev.mean")

@@

#check.gamViz(weighted_fl_abundance.2012.sqrt)
#weighted_fl_abundance.2012.sum.sqrt <- summary(weighted_fl_abundance.2012.sqrt)
#print(plot(weighted_fl_abundance.2012.sqrt, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2012.ck.sqrt <- check2D(weighted_fl_abundance.2012.sqrt, x1 = "yday", x2 = "elev.mean")

@@

#check.gamViz(weighted_fl_abundance.2012.bin)
#weighted_fl_abundance.2012.sum.bin <- summary(weighted_fl_abundance.2012.bin)
#print(plot(weighted_fl_abundance.2012.bin, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2012.ck.bin <- check2D(weighted_fl_abundance.2012.bin, x1 = "yday", x2 = "elev.mean")
```

@@

#check.gamViz(bb_abund.2010)
#bb_abund.2010.sum <- summary(bb_abund.2010)
#print(plot(bb_abund.2010, allTerms = TRUE), pages = 1)
#bb_abund.2010.ck <- check2D(bb_abund.2010, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals

@@

#check.gamViz(bb_abund.2011)
#bb_abund.2011.sum <- summary(bb_abund.2011)
#print(plot(bb_abund.2011, allTerms = TRUE), pages = 1)
#bb_abund.2011.ck <- check2D(bb_abund.2011, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals

@@

#check.gamViz(bb_abund.2012)
#bb_abund.2012.sum <- summary(bb_abund.2012)
#print(plot(bb_abund.2012, allTerms = TRUE), pages = 1)
#bb_abund.2012.ck <- check2D(bb_abund.2012, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals
```

@@

predict_pool.FL <- bind_rows(predict2010.FL, predict2011.FL, predict2012.FL) %>%
  mutate(date = as_date(yday),
         subject = factor(rep("FL", n())))
predict_pool.FL.sqrt <- bind_rows(predict2010.FL.sqrt, predict2011.FL.sqrt, predict2012.FL.sqrt) %>%
  mutate(date = as_date(yday),
         subject = factor(rep("FL", n())))
predict_pool.FL.bin <- bind_rows(predict2010.FL.bin, predict2011.FL.bin, predict2012.FL.bin) %>%
  mutate(date = as_date(yday),
         subject = factor(rep("FL", n())))

predict2010.BB <- predict_func(bb_abund.2010, newdata2010.BB, 2010)
predict2011.BB <- predict_func(bb_abund.2011, newdata2011.BB, 2011)
predict2012.BB <- predict_func(bb_abund.2012, newdata2012.BB, 2012)
predict_pool.BB <- bind_rows(predict2010.BB, predict2011.BB, predict2012.BB) %>%
  mutate(date = as_date(yday),
         subject = factor(rep("BB", n())),
         pred = log(pred))

predict_pool <- bind_rows(predict_pool.FL, predict_pool.BB)
predict_pool.sqrt <- bind_rows(predict_pool.FL.sqrt, predict_pool.BB)
predict_pool.bin <- bind_rows(predict_pool.FL.bin, predict_pool.BB)

# Subset predictions
predict_major <- filter(predict_pool, 
                           bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf")) %>%
  mutate(bb.sp = case_when(
    bb.sp == "bss" ~ "B. ss.",
    bb.sp == "pasc" ~ "B. pasc",
    bb.sp == "prat" ~ "B. prat",
    bb.sp == "soro" ~ "B. soro",
    bb.sp == "wurf" ~ "B. wurf"
  ))

predict_major.sqrt <- filter(predict_pool.sqrt, 
                           bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf")) %>%
  mutate(bb.sp = case_when(
    bb.sp == "bss" ~ "B. ss.",
    bb.sp == "pasc" ~ "B. pasc",
    bb.sp == "prat" ~ "B. prat",
    bb.sp == "soro" ~ "B. soro",
    bb.sp == "wurf" ~ "B. wurf"
  ))

predict_major.bin <- filter(predict_pool.bin, 
                           bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf")) %>%
  mutate(bb.sp = case_when(
    bb.sp == "bss" ~ "B. ss.",
    bb.sp == "pasc" ~ "B. pasc",
    bb.sp == "prat" ~ "B. prat",
    bb.sp == "soro" ~ "B. soro",
    bb.sp == "wurf" ~ "B. wurf"
  ))

predict_highelev <- filter(predict_pool, 
                              bb.sp %in% c("pyre", "mend", "mont")) %>%
  mutate(bb.sp = case_when(
    bb.sp == "pyre" ~ "B. pyre",
    bb.sp == "mend" ~ "B. mend",
    bb.sp == "mont" ~ "B. mont"
  ))

predict_highelev.sqrt <- filter(predict_pool.sqrt, 
                                   bb.sp %in% c("pyre", "mend", "mont")) %>%
  mutate(bb.sp = case_when(
    bb.sp == "pyre" ~ "B. pyre",
    bb.sp == "mend" ~ "B. mend",
    bb.sp == "mont" ~ "B. mont"
  ))

predict_highelev.bin <- filter(predict_pool.bin, 
                                  bb.sp %in% c("pyre", "mend", "mont")) %>%
  mutate(bb.sp = case_when(
    bb.sp == "pyre" ~ "B. pyre",
    bb.sp == "mend" ~ "B. mend",
    bb.sp == "mont" ~ "B. mont"
  ))

predict_minor <- filter(predict_pool, 
                           bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci")) %>%
  mutate(bb.sp = case_when(
    bb.sp == "gers" ~ "B. gers",
    bb.sp == "hort" ~ "B. hort",
    bb.sp == "hypn" ~ "B. hypn",
    bb.sp == "jone" ~ "B. jone",
    bb.sp == "lapi" ~ "B. lapi",
    bb.sp == "muci" ~ "B. muci"
  ))

predict_minor.sqrt <- filter(predict_pool.sqrt, 
                                bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci")) %>% 
  mutate(bb.sp = case_when(
    bb.sp == "gers" ~ "B. gers",
    bb.sp == "hort" ~ "B. hort",
    bb.sp == "hypn" ~ "B. hypn",
    bb.sp == "jone" ~ "B. jone",
    bb.sp == "lapi" ~ "B. lapi",
    bb.sp == "muci" ~ "B. muci"
  ))

predict_minor.bin <- filter(predict_pool.bin, 
                               bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci")) %>% 
  mutate(bb.sp = case_when(
    bb.sp == "gers" ~ "B. gers",
    bb.sp == "hort" ~ "B. hort",
    bb.sp == "hypn" ~ "B. hypn",
    bb.sp == "jone" ~ "B. jone",
    bb.sp == "lapi" ~ "B. lapi",
    bb.sp == "muci" ~ "B. muci"
  ))
```

### Visualize
```{r}
#### Major
ggplot(predict_major, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(abund.)") +
  #facet_grid(bb.sp + subject ~ year) +
  facet_grid(year ~ bb.sp + subject) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
ggsave("./output/GAM_major.png", height = 5, width = 11, units = "in")

ggplot(predict_major.sqrt, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(abund.)") +
  #facet_grid(bb.sp + subject ~ year) +
  facet_grid(year ~ bb.sp + subject) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
ggsave("./output/GAM_major_sqrt.png")

ggplot(predict_major.bin, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(abund.)") +
  #facet_grid(bb.sp + subject ~ year) +
  facet_grid(year ~ bb.sp + subject) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
ggsave("./output/GAM_major_bin.png")

#### High Elev
ggplot(predict_highelev, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(abund.)") +
  #facet_grid(bb.sp + subject ~ year) +
  facet_grid(year ~ bb.sp + subject) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
ggsave("./output/GAM_highelev.png", height = 5, width = 11, units = "in")

ggplot(predict_highelev.sqrt, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(abund.)") +
  #facet_grid(bb.sp + subject ~ year) +
  facet_grid(year ~ bb.sp + subject) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
ggsave("./output/GAM_highelev_sqrt.png", height = 5, width = 11, units = "in")

ggplot(predict_highelev.bin, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(abund.)") +
  #facet_grid(bb.sp + subject ~ year) +
  facet_grid(year ~ bb.sp + subject) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
ggsave("./output/GAM_highelev_bin.png", height = 5, width = 11, units = "in")

#### Minor
ggplot(predict_minor, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(abund.)") +
  #facet_grid(bb.sp + subject ~ year) +
  facet_grid(year ~ bb.sp + subject) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
ggsave("./output/GAM_minor.png", height = 5, width = 11, units = "in")

ggplot(predict_minor.sqrt, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(abund.)") +
  #facet_grid(bb.sp + subject ~ year) +
  facet_grid(year ~ bb.sp + subject) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
ggsave("./output/GAM_minor_sqrt.png", height = 5, width = 11, units = "in")

ggplot(predict_minor.bin, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(abund.)") +
  #facet_grid(bb.sp + subject ~ year) +
  facet_grid(year ~ bb.sp + subject) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
ggsave("./output/GAM_minor_bin.png", height = 5, width = 11, units = "in")
```
